Orienting the causal relationship between imprecisely measured traits using GWAS summary data

### 2017-11-13 docs/manuscript.Rmd
reorganising manuscript for submission

## Model

@@

First we describe how the CIT method [@Millstein2009] is implemented in the R package *R/cit* [@Millstein2016]. The methodology of the CIT is as follows. Assume an exposure $x$ is instrumented by a SNP $g$, and the exposure $x$ causes an outcome $y$, as described above. The following tests are then performed:

@@

The term in the 4th test can be rewritten as a $cov(g, y|x) = cov(g, y - \hat{y})$ where $y - \hat{y} = y - \hat{\alpha}_g + \hat{\beta}_g x$ is the residual of $y$ after adjusting for $x$, and $x$ is assumed to mediate the association between the SNP and the outcome. The condition in the 4th test is formulated as an equivalence testing problem that is estimated using simulations, comparing the estimate from the data against empirically obtained estimates for simulated variables where the independence model is true (full details are given in [@Millstein2009]). We note here that this approach is liable to fail, even when there is a true causal relationship, when confounders of the exposure and outcome are present, as these will induce collider bias.

@@

## Applied example using two sample MR

>>>>

### Model

@@

Here we describe how the CIT method [@Millstein2009] is implemented in the R package *R/cit* [@Millstein2016]. Assume an exposure $x$ is instrumented by a SNP $g$, and the exposure $x$ causes an outcome $y$, as described above. The following tests are then performed:

@@

The term in the 4th test can be rewritten as $cov(g, y|x) = cov(g, y - \hat{y})$ where $y - \hat{y} = y - \hat{\alpha}_g + \hat{\beta}_g x$ is the residual of $y$ after adjusting for $x$, and $x$ is assumed to mediate the association between the SNP and the outcome. The condition in the 4th test is formulated as an equivalence testing problem that is estimated using simulations, comparing the estimate from the data against empirically obtained estimates for simulated variables where the independence model is true (full details are given in [@Millstein2009]). We note here that this approach is liable to fail, even when there is a true causal relationship, when confounders of the exposure and outcome are present, as these will induce collider bias.

@@

### Applied example using two sample MR
